EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-06 draft.tex
Moved sections for the three experiments

% Clarify this is the first level.
In the case of the first level of integration, a decoupled worker, developers
can implement a new stateless docker container following the description of a
worker container described in the previous section (\ref{workers}). The worker
container executes a demon that continuously pulls messages from the input queue
on the Redis host specified as an environmental variable REDIS_HOST. If the new
population-based algorithm is written in python, developers can base the new
implementation on the provided worker container image, adding a new algorithm
function and worker and modifying the \texttt{main.py} file.
% say which function, what's its API, how can you test it - JJ
% this is not an API the need to write their own - Mario
% We can write an API later - M
If developers use an existing library for population-based algorithms, they need
to change the standard model of execution, that first creates a random
population because, in this case, we need to start with the population provided
as a parameter. Depending on the library, % Depending on which implementation? - JJ
it may be required to extract the population from the JSON message and replace
the population object created by the library. Again, the state of 
the population must be encoded as a JSON message to
be returned as a result. Additional data
must be returned in the JSON message, % how do you know the type of experiment? how do you add that? Which JSON key? - JJ
% I was wrong, it does not depend on the problem. - M
as depicted in  Listing \ref{code:result}. 
The output message contains the id of the worker container together
with records giving information about each iteration of the algorithm, 
including the best solution found. The controller uses this information 
to track the execution of the algorithm and keeping a log in the Redis 
container for posterior analyses. . % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
% Mandatory and used for everything - M

@@

% This is the second level? - JJ

On either level of integration, algorithm developers can upload the container
definition or images directly to a public repository like DockerHub. Another
option is to upload the container image definition (Dockerfile) to GitHub and
define a trigger to re-upload the image to the registry after an update. Having
images publicly available can contribute to the reproducibility of experiments.
Moreover, having more algorithms available can benefit researchers trying to
compare against other algorithms or increase the diversity of their algorithms
by adding more search methods. Finally, in some cloud services, having hosted 
images is a requirement, as it is the case of Amazon ECS.

% This is probably accessory. What difference would it make to upload it to GitHub?
% You also need to say in this section how it helps your overall objective. 
% Does it contribute to reproducibility? We've talked about that elsewhere. 
% Does it help diversity by making easy to add new algorithms? - JJ
% Explain it a bit more - M

@@

\subsection{Experiment Execution}

@@

\section{Experimental Setup} 
\label{setup}

We aimed to learn if the proposed solution can efficiently
improve the scalability and performance of population-based optimization
algorithms. Hence, in this section, we verify the following:

@@

\item Having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increases the performance of the
search by needing fewer function evaluations than a homogeneous setting?

\end{itemize}

@@

Next, we show the parameters for each algorithm, with Table \ref{tab:GAparams} for
the GA and Table \ref{tab:PSOparams} for PSO. We obtained these parameters
following the same method as in a previous work \cite{garcia2017benchmarking}.
To obtain the parameters, we tested first on the Rastrigin separable function
with five dimensions. After about fifteen experiments, the most challenging
targets were achieved for this particular function. We tested again with
functions one to three, and after obtaining favorable results, the PSO and GA
algorithm parameters were set. In the GA, we randomly set the mutation and
crossover probabilities to have more heterogeneous workers; % highlight this and add citation - JJ
in these parameters,
the range of values is specified. We did not change these parameters during the
experiments, and only the population size and number of generations were
provided as parameters.

@@

The experiments can be easily reproduced by using the configuration
files the the EvoSwarm application uses to run the noiseless BBOB testbed
experiments. % Specify this a bit more (with commands such as
             % docker-compose up and so on and use an itemize
             % environment to highlight it and answer to the reviewer
             % - JJ
As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. % Opening up the door to ask for testing with
                  % different proportion. You have to justify this
                  % value - JJ 
Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

% Section headers should be headlines. Like "Measuring scalability in
% heterogenous systems", or something like that. Experiment 1 says
% nothing about what's inside (this also applies below) - JJ
%\subsection{Experiment 1}
%\label{sec:exp1}
% Experiment 1

% What is the main objective of this experiment? Measure scalability?
% - JJ
As we mentioned before, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Other authors, like Salza et al., % add
                                % reference - JJ
have used a
similar architecture, which implies that worker nodes need to have at least a certain level
of complexity, in terms of execution time,  to effectively scale on multiple
nodes, tending to linear scalability. In a population-based algorithm like
EvoSwarm, several parameters can increase or decrease the workload of workers:

@@

@@

We executed two sets of experiments with five and ten initial populations,
repeating each experiment using  1, 2, 4, and 8 workers.  We kept the same
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list.

@@

%

% \subsection{Experiment 2}
% \label{sec:exp2}
% % Experiment 1

@@

\section{Experimental Results} 
\label{results}

We present the results of the experiments in this section. % that's pretty obvious. Say what you intend to prove with the experiments and why - JJ
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%

    \label{tab:medians}
    \vspace{0.25cm}
    \centering

\begin{table}[]
    \small
    \caption{Speedup achieved with incrementing the number of workers for each dimension showing the p-rate for the Wilcoxon test}
    % mark significant differences in some way - JJ
    \label{tab:medians}
    \vspace{0.25cm}
    \centering
    \begin{tabular}{lllllll}
              &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
    Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
    2         & 1w            & 2w            & 1.94            & 6E-16           & 1.84             & 2E-16           \\
              & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
              & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
    3         & 1w            & 2w            & 1.93            & 2E-16           & 1.90             & 2E-16           \\
              & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
              & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
    5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
              & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
              & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
    10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
              & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
              & 4w            & 8w            & 1.09            & 1E-03           & 1.78             & 2E-16           \\
    20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
              & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
              & 4w            & 8w            & 1.04            & 1E-02           & 1.80             & 2E-16           \\
    40        & 1w            & 2w            & 2.00            & 2E-16           & 1.95             & 2E-16           \\
              & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
              & 4w            & 8w            & 1.07            & 3E-02           & 1.98             & 2E-16          
    \end{tabular}
    \end{table}

\begin{figure}[h!p]
    \centering
    \includegraphics[width=\textwidth]{evalrate}
    \caption{Boxplot of the number of evaluations per second, for 30 instances of the function, with different number of workers and populations sizes }
    \label{fig:spworker}
\end{figure}

\begin{figure}[h!p]
    \centering
    \includegraphics[width=\textwidth]{evalsworker}
    \caption{Boxplot of the total number of evaluations for 30 instances of the function, with different number of workers and populations sizes }
    % Which function? - JJ
    \label{fig:spworker}
\end{figure}

%
%
% \subsection{Experiment 1}
% \label{sec:exp1results}
%
with results summarized in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. This is mainly
intended as a showcase of the flexibility of the tool presented in
this paper, but we also wanted to test how the number of workers
affected the time needed to reach a solution. Adding more populations
might add more overhead, so this might have a noticeable effect.
The left column in that figure has the
results when using five populations, and the right side shows the results for
ten. In principle the main issue is how to find the right number of
workers for a particular problem, but in general we wanted to check
the communication overhead when the size of the problem increases and
how the presented framework reacted to it. % Added some explanation
% here - JJ

\subsection{Deploying to a cloud provider}

>>>>

% Clarify this is the first level.
In the case of the first level of integration, a decoupled worker, developers
can implement a new stateless docker container following the description of a
worker container described in the previous section (\ref{workers}). The worker
container executes a demon that continuously pulls messages from the input queue
on the Redis host specified as an environmental variable {\texttt REDIS\_HOST}. If the new
population-based algorithm is written in python, developers can base the new
implementation on the provided worker container image, adding a new algorithm
function and worker and modifying the \texttt{main.py} file.
% say which function, what's its API, how can you test it - JJ
% this is not an API the need to write their own - Mario
% We can write an API later - M
If developers use an existing library for population-based algorithms, they need
to change the standard model of execution, that first creates a random
population because, in this case, we need to start with the population provided
as a parameter. Depending on the library, % Depending on which implementation? - JJ
it may be required to extract the population from the JSON message and replace
the population object created by the library. Again, the state of 
the population must be encoded as a JSON message to
be returned as a result. Additional data
must be returned in the JSON message, % how do you know the type of experiment? how do you add that? Which JSON key? - JJ
% I was wrong, it does not depend on the problem. - M
as depicted in  Listing \ref{code:result}. 
The output message contains the id of the worker container together
with records giving information about each iteration of the algorithm, 
including the best solution found. The controller uses this information 
to track the execution of the algorithm and keeping a log in the Redis 
container for posterior analyses.  % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
% Mandatory and used for everything - M

@@

% This is the second level? - JJ
% either - M
On either level of integration, algorithm developers can upload the container
definition or images directly to a public repository like DockerHub. Another
option is to upload the container image definition (Dockerfile) to GitHub and
define a trigger to re-upload the image to the registry after an update. Having
images publicly available can contribute to the reproducibility of experiments.
Moreover, having more algorithms available can benefit researchers trying to
compare against other algorithms or increase the diversity of their algorithms
by adding more search methods. Finally, in some cloud services, having hosted 
images is a requirement, as it is the case of Amazon ECS.

% This is probably accessory. What difference would it make to upload it to GitHub?
% You also need to say in this section how it helps your overall objective. 
% Does it contribute to reproducibility? We've talked about that elsewhere. 
% Does it help diversity by making easy to add new algorithms? - JJ
% I explain it a bit more - M

@@

\subsection{Experiment Execution}
\label{sec:exeperimentalDefinition}

@@

\section{General Experimental Setup} 
\label{setup}

We aimed to learn if the proposed solution can efficiently
improve the scalability and performance of population-based optimization
algorithms. Hence, in the following sections, we verify the following:

@@

\item Having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increases the performance of the
search by needing fewer function evaluations than a homogeneous setting?
\end{itemize}

@@

Next, we show the parameters for each algorithm, with Table \ref{tab:GAparams} for
the GA and Table \ref{tab:PSOparams} for PSO. We obtained these parameters
following the same method as in a previous work \cite{garcia2017benchmarking}.
To obtain the parameters, we tested first on the Rastrigin separable function
with five dimensions. After about fifteen experiments, the most challenging
targets were achieved for this particular function. We tested again with
functions one to three, and after obtaining favorable results, the PSO and GA
algorithm parameters were set. In the GA, we randomly set the mutation and
crossover probabilities to have more heterogeneous workers; % highlight this and add citation - JJ 
% done below - M
in these parameters, the range of values is specified. We did not change these parameters during the 
experiments, and only the population size and number of generations were
provided as parameters. This strategy has been applied successfully for increasing 
the performance of multi-population algorithms without requiring the setting 
of initial parameters \cite{garcia2014randomized}.

@@

The experiments can be easily reproduced by using the configuration
files the the EvoSwarm application uses to run the noiseless BBOB testbed
experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
             % docker-compose up and so on and use an itemize
             % environment to highlight it and answer to the reviewer
             % - JJ
As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. % Opening up the door to ask for testing with
                  % different proportion. You have to justify this
                  % value - JJ  
                  % We only tested 0, 1 and .5, there is not really a justification - M 
Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

% Section headers should be headlines. Like "Measuring scalability in
% heterogenous systems", or something like that. Experiment 1 says
% nothing about what's inside (this also applies below) - JJ
\section{Measuring Scalability}
\label{sec:exp1}
% Experiment 1
As we mentioned before, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},   % add
                                % reference - JJ
have used a
similar architecture, which implies that worker nodes need to have at least a certain level
of complexity, in terms of execution time,  to effectively scale on multiple
nodes, tending to linear scalability. In a population-based algorithm like
EvoSwarm, several parameters can increase or decrease the workload of workers:

@@

% What is the main objective of this experiment? Measure scalability?
% - JJ 
% Yes, see below - M

@@

\subsection{Setup}

We executed two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list.

with results summarized in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. This is mainly
intended as a showcase of the flexibility of the tool presented in
this paper, but we also wanted to test how the number of workers
affected the time needed to reach a solution. Adding more populations
might add more overhead, so this might have a noticeable effect.
The left column in that figure has the
results when using five populations, and the right side shows the results for
ten. In principle the main issue is how to find the right number of
workers for a particular problem, but in general we wanted to check
the communication overhead when the size of the problem increases and
how the presented framework reacted to it. % Added some explanation
% here - JJ

@@

\subsection{Experimental Results} 
\label{results}
% that's pretty obvious. Say what you intend to prove with the experiments and why - JJ 
% Removed - M
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%

    \label{tab:medians:rate}
    \vspace{0.25cm}
    \centering

\begin{table}[]
    \small
    \caption{Speedup achieved with incrementing the number of workers for each dimension showing the p-rate for the Wilcoxon test}
    % mark significant differences in some way - JJ
    \label{tab:speedup:test}
    \vspace{0.25cm}
    \centering
    \begin{tabular}{lllllll}
              &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
    Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
    2         & 1w            & 2w            & 1.94            & 6E-16           & 1.84             & 2E-16           \\
              & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
              & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
    3         & 1w            & 2w            & 1.93            & 2E-16           & 1.90             & 2E-16           \\
              & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
              & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
    5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
              & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
              & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
    10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
              & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
              & 4w            & 8w            & 1.09            & 1E-03           & 1.78             & 2E-16           \\
    20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
              & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
              & 4w            & 8w            & 1.04            & 1E-02           & 1.80             & 2E-16           \\
    40        & 1w            & 2w            & 2.00            & 2E-16           & 1.95             & 2E-16           \\
              & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
              & 4w            & 8w            & 1.07            & 3E-02           & 1.98             & 2E-16          
    \end{tabular}
    \end{table}

\begin{figure}[h!p]
    \centering
    \includegraphics[width=\textwidth]{evalrate}
    \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
    \label{fig:spworker:ratio}
\end{figure}

\begin{figure}[h!p]
    \centering
    \includegraphics[width=\textwidth]{evalsworker}
    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
    % Which function? - JJ
    \label{fig:spworker}
\end{figure}

\section{Performance of Heterogeneous Populations in the BBOB Benchmark}
\label{sec:exp2}

\subsection{Setup}

% % Experiment 1

@@

\section{Deploying to a cloud provider}
